Predicting computational reproducibility of data analysis pipelines \\in large population studies using collaborative filtering

### 2018-07-26 paper.tex
A few fixes and comments

\paragraph{Random File Numbers (Triangular) -- RFNT}
The number of files selected for every subject is randomly selected in
a triangular distribution $T(a, b, c)$ as in Figure~\ref{fig:triangular}. The mean of the distribution is 
$\frac{a+b+c}{3}$. We set \textit{c} to $N_{f}$ and we set 
\textit{a} and \textit{b} with two approaches that ensure that the 
average of the distribution is $\alpha N_{f}$:
\begin{enumerate}
        \item Largest a (RFNTL, Figure~\ref{fig:triangular-L-Sample-Training-set}): a is set to the 
        largest possible value, i.e., b. The average of the 
        distribution is $\frac{2a+N_{f}}{3}$, therefore 
        $a=\frac{3\alpha-1}{2}N_{f}$. When $\alpha \leq 1/3$, the 
        number of files in every subject is sampled from $T(0, 0, N_f)$ 
        with probability $3\alpha$ and set to 0 otherwise, to ensure 
        that the average number of files by subject in the training set 
        is $\alpha N_f$.
        \item Smallest a (RFNTS, Figure~\ref{fig:triangular-S-Sample-Training-set}): a is 
        set to the smallest possible value, i.e., 0. The average of the 
        distribution is $\frac{b+N_{f}}{3}$, therefore 
        $b=N_{f}(3\alpha-1)$. This implies that $\alpha \geq 1/3$, so 
        that $b \geq 0$, and $\alpha \leq 2/3$, so that $b \leq N_f$. 
        When $\alpha < 1/3$, we set $b=0$ and the number of files 
        selected in every subject is sampled from $T(0, 0, N_f)$ with 
        probability $3\alpha$ and set to 0 otherwise, as in RFNTL. 
        When $\alpha \geq 2/3$, we set $b=N_f$ and the number of files 
        selected in every subject is sampled from $T(0, N_f, N_f)$ with 
        probability $3(1-\alpha)$ and set to $N_f$ otherwise, still to 
        ensure that the average number of files by subject in the training set is 
        $\alpha N_f$.
\end{enumerate}
As illustrated in Figures~\ref{fig:triangular-L-Sample-Training-set} 
and~\ref{fig:triangular-S-Sample-Training-set}, the motivation for the 
RFNTL method is to guarantee that, for large enough values of 
$\alpha$, all subjects will have at least a few files in the training 
set, which is not the case for RFNU or RFNTS.
\begin{figure}
\centering
\includegraphics[width=0.5\columnwidth]{figures/triangular.pdf}
\caption{Triangular distribution $T(a, b, c)$}
\label{fig:triangular}
\end{figure}

@@

Two experiments have been conducted for each reproducibility matrix, to evaluate the performance of our 
predictions using (1) ALS without biases, and (2) ALS with biases. 
Results will be evaluated using classification accuracy and compared to (1) a dummy classifier that always 
predicts the value in the majority class and (2) the Random Unreal 
method, used as the baseline sampling technique. All reported values are averages
over 5 repetitions. Due to sampling issues, it is possible that the actual 
training ratio obtained with some of the sampling methods does not 
match the target one. We checked that the difference between the target 
and real training ratios was lower than 0.01. We used Spark's ALS model 
as available in package \texttt{pyspark.ml}, with 50 factors, a 
regularization parameter of 0.01, a maximum of 5 iterations and non 
negative constraints set to true.

@@

Accuracy results for the different numbers of subject types are 
reported in Figure~\ref{fig:results-synthetic}. By construction, the 
accuracy of the dummy classifier is close to 0.5 for all subject types \todo{Why such variations?}. Random 
Unreal performs very well for all subject types, which confirms that 
ALS is working correctly. Unsurprisingly, all the other methods perform 
better than the dummy classifier, and their accuracy decreases as the 
number of subject types increases.

@@

RS, RFNU and RFNTL achieve a 90\% accuracy for a training ratio of 
about 0.85. Therefore, 15\% of the computations could potentially be 
saved in such reproducibility studies.

@@

\todo{Different number of factors.}

@@

@@

\todo{Discussion on points that are not part of constant lines.}

\todo{Show an overlap to see where the error is located.}

\todo{Make sure the code is available and results are reproducible.}

>>>>

\paragraph{Random File Numbers (Triangular) -- RFNT}
The number of files selected for every subject is randomly selected in
a triangular distribution $T(a, b, c)$ as in Figure~\ref{fig:triangular}. The mean of the distribution is 
$\frac{a+b+c}{3}$. We set \textit{c} to $N_{f}$ and we set 
\textit{a} and \textit{b} with two approaches that ensure that the 
average of the distribution is $\alpha N_{f}$: \todo{clarify this part}
\begin{enumerate}
        \item Largest a (RFNTL, Figure~\ref{fig:triangular-L-Sample-Training-set}): a is set to the 
        largest possible value, i.e., b. The average of the 
        distribution is $\frac{2a+N_{f}}{3}$, therefore 
        $a=\frac{3\alpha-1}{2}N_{f}$. When $\alpha \leq 1/3$, the 
        number of files in every subject is sampled from $T(0, 0, N_f)$ 
        with probability $3\alpha$ and set to 0 otherwise, to ensure 
        that the average number of files by subject in the training set 
        is $\alpha N_f$.
        \item Smallest a (RFNTS, Figure~\ref{fig:triangular-S-Sample-Training-set}): a is 
        set to the smallest possible value, i.e., 0. The average of the 
        distribution is $\frac{b+N_{f}}{3}$, therefore 
        $b=N_{f}(3\alpha-1)$. This implies that $\alpha \geq 1/3$, so 
        that $b \geq 0$, and $\alpha \leq 2/3$, so that $b \leq N_f$. 
        When $\alpha < 1/3$, we set $b=0$ and the number of files 
        selected in every subject is sampled from $T(0, 0, N_f)$ with 
        probability $3\alpha$ and set to 0 otherwise, as in RFNTL. 
        When $\alpha \geq 2/3$, we set $b=N_f$ and the number of files 
        selected in every subject is sampled from $T(0, N_f, N_f)$ with 
        probability $3(1-\alpha)$ and set to $N_f$ otherwise, still to 
        ensure that the average number of files by subject in the training set is 
        $\alpha N_f$.
\end{enumerate}
As illustrated in Figures~\ref{fig:triangular-L-Sample-Training-set} 
and~\ref{fig:triangular-S-Sample-Training-set}, the motivation for the 
RFNTL method is to guarantee that, for large enough values of 
$\alpha$, all subjects will have at least a few files in the training 
set, which is not the case for RFNU or RFNTS.
\begin{figure}
\centering
\includegraphics[width=0.5\columnwidth]{figures/triangular.pdf}
\caption{Triangular distribution $T(a, b, c)$}
\label{fig:triangular}
\end{figure}

@@

Two experiments have been conducted for each reproducibility matrix, to evaluate the performance of our 
predictions using (1) ALS without biases, and (2) ALS with biases. 
Results will be evaluated using classification accuracy and compared to (1) a dummy classifier that always 
predicts the value in the majority class and (2) the Random Unreal 
method, used as the baseline sampling technique. All reported values are averages
over 5 repetitions. Due to sampling issues, it is possible that the actual 
training ratio obtained with some of the sampling methods does not 
match the target one. We checked that the difference between the target 
and real training ratios was lower than 0.01. We used Spark's ALS model 
as available in package \texttt{pyspark.ml}, with 50 factors, a 
regularization parameter of 0.01, a maximum of 5 iterations and non 
negative constraints set to true. \todo{Tristan: Add one plot with sensitivity and specificity for 0.9 training}

@@

Accuracy results for the different numbers of subject types are 
reported in Figure~\ref{fig:results-synthetic}. By construction, the 
accuracy of the dummy classifier is close to 0.5 for all subject types \todo{Soudabeh: why such variations?}. Random 
Unreal performs very well for all subject types, which confirms that 
ALS is working correctly. Unsurprisingly, all the other methods perform 
better than the dummy classifier, and their accuracy decreases as the 
number of subject types increases.

@@

RS, RFNU and RFNTL achieve a 90\% accuracy for a training ratio of 
about 0.85 \todo{precise that}. Therefore, 15\% of the computations could potentially be 
saved in such reproducibility studies.

@@

\todo{Soudabeh: Different number of factors: type 8, RFNU, 50, 100, 200, 400.}

@@

Figure~\ref{fig:overlap} illustrates the comparison between RFNU, RFNTL 
and RFNTS. The matrices in this Figure represent the union between the 
training and the test set, where the training set is represented in 
black (negatives) and white (positives), and the test set is 
represented in green (true positives), yellow (false negatives), gray 
(true negatives) and red (false positives). This representation 
provides insights regarding where, and perhaps why, prediction errors 
occur. At this training ratio, RFNU (Figure~\ref{fig:overlap-rfntu}) 
uniformly samples the number of files per subject between 0.8$N_f$ and 
$N_f$. It provides an opportunity for ALS to be trained on the last 
files of the pipelines, while maintaining a low number of columns with 
a low training ratio. On the contrary, RFNTL 
(Figure~\ref{fig:overlap-rfntl}) samples the number of files per 
subject from 0.85$N_f$, but the probability to have a complete column 
is 0 (the one complete column in Figure~\ref{fig:overlap-rfntl} is the one included
to prevent cold start issues).

\begin{figure*}
\begin{subfigure}[b]{0.6\columnwidth}
        \includegraphics[width=\columnwidth]{figures/RFNU_Syn_8_ALS_09_test_data_matrix_run1.png}
        \caption{RFNU: $U(0.8N_f, N_f)$}
        \label{fig:overlap-rfnu}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.6\columnwidth}
        \includegraphics[width=\columnwidth]{figures/RFNT-L_Syn_8_ALS_09_test_data_matrix_run2.png}
        \caption{RFNTL: $T(0.85N_f, 0.85N_f, N_f)$}
        \label{fig:overlap-rfntl}
\end{subfigure}
\hfill
\begin{subfigure}[b]{0.6\columnwidth}
        \includegraphics[width=\columnwidth]{figures/RFNT-S_Syn_8_ALS_09_test_data_matrix_run1.png}
        \caption{RFNTS: 0.3$T(0, N_f, N_f)$+0.7$N_f$}
        \label{fig:overlap-rfnts}
\end{subfigure}
\caption{Comparison between RFNU, RFNTL and RFNTS on synthetic dataset with 8 types, $\alpha=0.9$.}
\label{fig:overlap}
\end{figure*}

@@

\todo{Discussion on points that are not part of constant lines. Soudabeh: add an overlap matrix for FS, RFNU, ALS, at 0.5 or 0.6.}

\todo{Make sure the code is available and results are reproducible. Soudabeh: push your code to reprotools. Tristan: release and check.}
